\section{Concluding remarks}
\label{sec:conclude}

This paper presents efficient formulas to compute Miller doubling and Miller addition on curves of $j$-invariant 0 with embedding degree divisible by 3 when written in Hessian form.
This paper presents formulas for both pairings of the form $\G{1} \times \G{2}$ and $\G{2} \times \G{1}$ and compares the efficiency of these formulas to the best known formulas of previous research.
We present the first formulas for pairings on $\G{2} \times \G{1}$ that utilize twists of degree 3 in the case of odd embedding degrees, and the first formulas that utilize twists of degree 3 for Hessian curves in all cases. Our tables suggest that, for embedding degrees divisible by 6, the most efficient choices from the known implementations (including those in this paper) are degree 6 twists of Edwards curves for pairings on $\G{1} \times \G{2}$ and degree 6 twists of Weierstrass curves with projective coordinates for pairings on $\G{2} \times \G{1}$.
For embedding degrees divisible by 3 but not 6, the most efficient choices from the 
known implementations (including those is in this paper) are degree 3 twists of Weierstrass curves with projective coordinates for pairings on $\G{1} \times \G{2}$ 
and degree 3 twists of Hessian curves (that is, the constructions from this paper) for pairings on $\G{2} \times \G{1}$. 
Furthermore, for embedding degrees 15 and 21 we expect the optimal ate pairing with Hessian curves to be the fastest choice; this outperforms
the twisted ate pairing (the best choice for $\G{1} \times \G{2}$) with projective coordinates of Weierstrass curves by a factor of between 3 and 4, as explained in the previous section.

Curves generated by the methods used in this paper (originally due to~\cite{2010/freeman}) are guaranteed to have twists of degree $3$
and have embedding degree
$k \equiv 3,9,15 \pmod{18}$ or
$k \equiv 0 \pmod{6}$ where $18 \nmid k$.
As explained in earlier sections, we suggest updating the use of embedding degree 12 to 15 for 128-bit security and 18 to 21 for 192-bit security in light of the NFS attacks and their variants. This allows us to keep the relatively small primes for the base field and the low $\rho$-value of the families described above. The added security comes from the fact that 15 and 21 are slightly larger than 12 and 18 respectively, and both have only 2 prime factors, hence are the cases for which the NFS techniques are least effective. 
As both of these embedding degrees are odd, we suggest the use of Hessian curves with degree 3 twists in combination with the optimal ate pairing. 
This the most efficient available implementation choice in this context -- as explained in the discussion of the previous section.

In future work,
we plan to study precisely how the NFS attacks and their variants apply to our constructions in order to be able to
properly evaluate the security and propose concrete parameters.
A trade-off between the low $\rho$-value and large, odd embedding degrees that we suggest in this paper and the higher $\rho$-value and small embedding degrees suggested in~\cite{2018/FK} would be very interesting, but we leave this for future work.
It would also be interesting to evaluate the performance of other curve models with degree 3 twists on $\G{2} \times \G{1}$ pairings.
We also consider the optimized implementation as future work.

%Our formulas for even embedding degrees are the fastest among all the previously published one. - No they're not (twisted Edwards are faster)
%We also describe techniques to eliminate intermediate denominators for odd embedding degrees. - this isn't new

%Although pairings on Hessian curves have already been considered,
%for example, by Gu, Gu, and Xie in~\cite{2010/Gu} and by Li and Zhang in~\cite{2012/Li},
%no methods regarding how to generate pairing-friendly Hessian curves were described.

%The pairing-friendly families of BN, BLS, and KSS curves use primarily embedding degrees 12 or 18.
%Due to recent advances in the discrete logarithm problem, it has become necessary to increase the size of the finite field,
%which can be done either by increasing the size of the prime or by increasing the embedding degree.
%Our constructions address this problem by generating pairing-friendly twisted Hessian curves with embedding degrees 15 and 21,
%which allows for an increase in security without having to jump from embedding degree 12 to 18, or from 18 to 24.

%%XXX emphasize advantages not shown in the tables
%There are various advantages of our constructions that cannot be justified by the cost shown in Table~\ref{tbl-cmp}. %Table~\ref{tbl-even} and Table~\ref{tbl-odd}.
%First of all, the costs shown in those tables do not reflect the total cost for the pairing-based protocols.
%This is because generically there are many group operations performed prior to relatively few pairing computations,
%and group operations on twisted Hessian curves allow faster point arithmetic operations than Weierstrass curves.
%Secondly, having curves represented in the same models for $\mathbb{G}_1$ and $\mathbb{G}_2$ does not induce the extra cost of conversion between curve models.
%(Recall that for BN, BLS, and KSS, this conversion is always necessary if one wants to take advantage of
%the fast point-arithmetic on Hessian or Edwards curves, as proven in~\cite{2013/bos-pairing}.)

%We also would like to emphasize that embedding degrees such as 12, 18, 24 can split into many subfields.
%Recall that recent advances in number field sieve attacks on solving DLP~\cite{2016/KB} which further generalized by~\cite{2017/KJ} and~\cite{2016/conjugate}
%%~\cite{2016/SS}
%target on finite field $\F{q^n}$ where $q$ is prime and $n$ is composite.
%This implies that the attacks apply to pairings which use composite extension degrees.
%
%%XXX Laurent said not convincing !!!
%To be more precise, these attacks rely on splitting $n$ into $n = \eta \kappa$ for non-trivial factorization.
%Since, for example, 12 can split as $(\eta,\kappa) = (2,6), (3,4), (4,3), (6,2)$,
%it allows many possibles choices of subfields to perform the attacks.
%Similar remarks also apply to
%$18$ which can be split as $(2,9), (3,6), (6,3), (9,2)$
%and
%$k = 24$ which can be split as $(2,12)$, $(3,8)$, $(4,6)$, $(6,4)$, $(8,3)$, $(12,2)$.
%%Note that some methods above all so have a restriction of $gcd(\eta,\kappa) = 1$, but some of the do not force this condition.
%
%On the other hand, there are sufficiently fewer subfields in our case.
%For example, 15 splits into either $(3,5)$ or $(5,3)$
%while 21 splits into either $(3,7)$ or $(7,3)$.
%This means that we reduce the possible choices of subfields that the attacks can target.

%In this paper, we do not propose any concrete parameters for our pairing-friendly twisted Hessian curves.
%This is due to the fact that the complexity of the new attacks have neither been completely understood nor reached the stable stage yet.
%Moreover, there is no analysis of how those attacks would apply to embedding degree 15 or 21.

